---
title: "Diagnostics"
author: "Fabian Gunzinger"
date: "1/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(plm)
library(fixest)
library(ggplot2)
library(gglm)
library(ggfortify)
library(MatrixModels)
library(patchwork)
source('helpers.R')
setwd('~/dev/projects/entropy/entropy/analysis')
FIGDIR = '/Users/fgu/dev/projects/entropy/output/figures'
```


Below we perform a set of diagnostics to test the assumptions on our main specification.

```{r}
dt <- read_analysis_data()
data <- dt[sa_inflows <= 3000 & sa_outflows <= 3000]
```


```{r}
controls = c(
  # fin behaviour
  'pct_credit',
  'month_spend',
  # planning - tbd
  # hh / ind chars
  'is_urban',
  'month_income',
  'has_month_income',
  'income_var',
  'has_rent_pmt',
  'has_mortgage_pmt',
  'has_loan_repmt',
  'has_benefits'
)

# estimated without fixed-effects so we can calculate hat values (for leverage)
# see: https://lrberge.github.io/fixest/reference/hatvalues.fixest.html
# m <- fixest::feols(xpd(has_sa_inflows ~ entropy_tag_spend_s + ..controls + factor(user_id) + factor(ym), ..controls = controls), data=data)

# faster fe version, in case hat values not needed
m <- fixest::feols(xpd(sa_inflows ~ entropy_tag_spend_s + ..controls +factor(user_id) + factor(ym), ..controls = controls), data=data)
```



```{r}
d <- data.table(
  fitted = fitted(m),
  resid = resid(m),
  std_resid = resid / sigma(m),
  sqrt_std_resid = sqrt(abs(std_resid))
)

resid_fitted <- ggplot(d, aes(fitted, resid)) +
  geom_point(aes(colour = factor(data$has_sa_inflows))) +
  labs(
    title = 'Residuals vs fitted',
    x = 'Fitted',
    y = 'Residuals',
    colour = 'Has sa inflows'
  )
resid_fitted


qq <- ggplot(d, aes(sample = std_resid)) + 
  geom_qq_line(linetype = "dashed") +
  stat_qq(alpha = 0.5) +
  labs(
    title = "Normal Q-Q",
    x = "Theoretical Quantiles",
    y = "Standardised Residuals"
  )


scale_loc <- ggplot(d, aes(fitted, sqrt_std_resid)) +
  geom_point() +
  geom_smooth() +
  labs(
    title = 'Scale-location',
    x = 'Fitted values',
    y = 'Sqrt of standardised residuals'
  )

resid_leverage <- ggplot() +
  geom_point(aes(std_resid, hatvalues(m))) +
  geom_smooth(aes(std_resid, hatvalues(m))) +
  labs(
    title = 'Residual vs legerage',
    x = 'Leverage',
    y = 'Standardised residuals'
  )

pw <- resid_fitted + qq + scale_loc
pw
# ggsave(file.path(FIGDIR, 'diagnostics.png'))
```


## Model assumptions

Our model is estimated under the following assumptions:

- Linear in parameters
- Homoskedasticity
- Independence
- Normality (y is normally distributed for fixed x)


Interpretation:

- Residual vs fitted plot helps check for linearity and homoskedasticity: under linearity, we'd expect residuals to not be too large (as they would be if, for instance, we were to fit a straight line to onto quadratic data), under homoskedasticity we'd expect the size of the residuals to be independent of the fitted values.

- QQ-plot helps check for normality: diagonal line represent ideal normal observations, so the more closely the model residuals lie along that line, the better the assumption of normality holds.

- The scale location plot helps check for homoskedasticity: if the assumption is valid, we expect no pattern (just as in the residual vs fitted plot).

- The residual-leverage plot helps us detect outliers with respect to the independent variables. (Taks a long time to calculate, so not currently done.)


```{r}
resid
```



## Useful sources

- https://bookdown.org/dereksonderegger/571/7-Diagnostics-Chapter.html

- http://karthur.org/2019/implementing-fixed-effects-panel-models-in-r.html

- https://github.com/graysonwhite/gglm
