---
title: "Savings behaviour"
author: "Fabian Gunzinger"
date: "16/03/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('~/dev/projects/entropy/entropy/analysis')
source('helpers.R')

library(ggplot2)
library(lubridate)
library(patchwork)


FIGDIR = '/Users/fgu/dev/projects/entropy/output/figures' 
SAMPLE = 'X77'

# load data
dt = read_txn_data(SAMPLE)
```

Create savings data for users in final sample (as well as a dataset with key variables for convenient inspection).

```{r}
# savings txns of users in final sample
dta = read_analysis_data(SAMPLE)
final_sample <- unique(dta$user_id)
dt <- dt[user_id %in% final_sample,]
savings <- dt[is_sa_flow == 1 & is_debit == FALSE]
setorderv(savings, c('user_id', 'date'))

```


## Savings frequency

How frequently do users transfer money into their savings accounts?

```{r}
savings[, date_diff := difftime(date, shift(date), units = 'days'), .(user_id)]
savings <- savings[, .SD[date_diff < quantile(date_diff, probs = .95, na.rm = T)]]
setcolorder(savings, c('user_id', 'date', 'account_id', 'date_diff'))

p1 <- ggplot(savings) + 
  geom_bar(aes(factor(date_diff)))
```

The pattern is roughly what we'd expect, with spikes on a weekly, biweekly, and monthly basis.

What is puzzling is the rather high number of txns that happen less than a week after the last one, and the large number of txns that happen on the same day.

Accounting for transfers to multiple accounts.

```{r}
savings[, date_diff2 := difftime(date, shift(date), units = 'days'), .(user_id, account_id)]
savings <- savings[, .SD[date_diff2 < quantile(date_diff2, probs = .95, na.rm = T)]]
setcolorder(savings, c('user_id', 'date', 'account_id', 'date_diff', 'date_diff2'))

p2 <- ggplot(savings) + 
  geom_bar(aes(factor(date_diff2)))
```

Accounting for potential duplicates (txns of same amount)

```{r}
savings[, date_diff3 := difftime(date, shift(date), units = 'days'), .(user_id, account_id, amount)]
savings <- savings[, .SD[date_diff3 < quantile(date_diff3, probs = .95, na.rm = T)]]
setcolorder(savings, c('user_id', 'date', 'account_id', 'date_diff', 'date_diff2', 'date_diff3'))

p3 <- ggplot(savings) + 
  geom_bar(aes(factor(date_diff3)))
```



```{r}
p1 + p2 + p3

```


Regular savings txns of same of same amount and into same account happen, as we'd expect, either on a weekly or monthly basis.
```{r}
ggplot(savings) +
  geom_bar(aes(date_diff3)) +
  xlim(0, 40)
```



```{r}
# view data for a random user
users <- unique(savings$user_id)
```
```{r}
user <- sample(users, size = 1)
dd[user_id == user]
```









Inspection reveals that often, savings txs on the same day go into different accounts, which makes sense. For txns with a diff of one or two days, I suspect some of these are also duplicates that we aren't able to detect.

Inspecting also reveals some clear differences in savings patterns among users: some users save small amount weekly, some save monthly, some make transfers to multiple accounts on a regular basis (often monthly), some save without any clear pattern (possibly whenever they have money, time, or remember to save). **This is something to explore further. Could possibly try to infer clusters of savings types.**


```{r}
summary(savings)
```


## Including Plots

You can also embed plots, for example:

```{r}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
